Temporal ordering and registration of images in studies of developmental dynamics

### 2014-09-03 image_analysis_paper2/SI_text.tex
changed tex to have bib things in tex for PNAs submission

\bibliographystyle{pnas2}
\bibliography{../image_analysis_paper/background_reading/references,../../references/references}

>>>>

%\bibliographystyle{pnas2}
%\bibliography{../image_analysis_paper/background_reading/references,../../references/references}

\begin{thebibliography}{10}

\bibitem{canny1986computational}
Canny J (1986) A computational approach to edge detection. \textit{IEEE Trans Pattern Anal Mach Intell} (6):679--698.

\bibitem{singer2011angular}
Singer A (2011) Angular synchronization by eigenvectors and semidefinite
  programming. \textit{Appl Comput Harmon Anal}
  30(1):20--36.

\bibitem{ahuja2007template}
Ahuja S, Kevrekidis I~G, Rowley C~W (2007) Template-based stabilization of
  relative equilibria in systems with continuous symmetry. \textit{Journal of
  Nonlinear Science} 17(2):109--143.

\bibitem{goemans1995improved}
Goemans M~X, Williamson D~P (1995) Improved approximation algorithms for
  maximum cut and satisfiability problems using semidefinite programming.
  \textit{Journal of the ACM (JACM)} 42(6):1115--1145.

\bibitem{bandeira2013cheeger}
Bandeira A~S, Singer A, Spielman D~A (2013) A {C}heeger inequality for the
  graph connection {L}aplacian. \textit{SIAM Journal on Matrix Analysis and
  Applications} 34(4):1611--1630.

\bibitem{coifman2005geometric}
Coifman R~R, Lafon S, Lee A~B, Maggioni M, Nadler B, \textit{et~al.} (2005)
  Geometric diffusions as a tool for harmonic analysis and structure definition
  of data: Diffusion maps. \textit{Proc Natl Acad Sci USA} 102(21):7426--7431.

\bibitem{coifman2008graph}
Coifman R~R, Shkolnisky Y, Sigworth F~J, Singer A (2008) Graph {L}aplacian
  tomography from unknown random projections. \textit{IEEE Trans Image Process} 17(10):1891--1899.

\bibitem{rohrdanz2011determination}
Rohrdanz M~A, Zheng W, Maggioni M, Clementi C (2011) Determination of reaction
  coordinates via locally scaled diffusion map. \textit{The J Chem Phys} 134:124116.

\bibitem{Belkin2003}
Belkin M, Niyogi P (2003) Laplacian eigenmaps for dimensionality reduction and
  data representation. \textit{Neural Comput} 15:1373--1396.

\bibitem{singer2012vector}
Singer A, Wu H~T (2012) Vector diffusion maps and the connection {L}aplacian.
  \textit{Commun Pure Appl Math} 65(8):1067--1144.

\bibitem{ferguson2010systematic}
Ferguson A~L, Panagiotopoulos A~Z, Debenedetti P~G, Kevrekidis I~G (2010)
  Systematic determination of order parameters for chain dynamics using
  diffusion maps. \textit{Proc Natl Acad Sci USA}
  107(31):13597--13602.

\bibitem{kemelmacher2011exploring}
Kemelmacher-Shlizerman I, Shechtman E, Garg R, Seitz S~M (2011) Exploring
  photobios, \textit{ACM Trans Graph} (ACM), volume~30,
  p.~61.

\end{thebibliography}

### 2014-09-03 image_analysis_paper2/pnas_draft.tex
everything updated for resubmission
changed tex to have bib things in tex for PNAs submission

\graphicspath{ {paper_figures2/} }

@@

@@

\significancetext{
Spatial registration and temporal ordering of images of developing embryos are one of the first steps in the reconstruction of developmental dynamics. 
%
When the number of images is small and the differences between them are easily recognizable, image registration and ordering can be done manually; these tasks become challenging when the data sets are larger and more complex. 
%
Here we show this can be automated and combined in a single
step using recent advances in data mining. 
%
To illustrate this approach, we order and register data from imaging studies of {\it Drosophila}embryogenesis. }

@@

@@

@@

@@

\bibliographystyle{pnas2}
\bibliography{../../references/references,../image_analysis_paper/background_reading/references}

>>>>

@@

\font\xfigtextfontit=\frutigeroblique at 7pt

@@

\significancetext{
Spatial registration and temporal ordering of images of developing embryos are one of the first steps in the reconstruction of developmental dynamics. 
%
When the number of images is small and the differences between them are easily recognizable, image registration and ordering can be done manually; these tasks become challenging when the data sets are larger and more complex. 
%
Here we show this can be automated and combined in a single
step using recent advances in data mining. 
%
To illustrate this approach, we order and register data from imaging studies of {\xfigtextfontit  Drosophila} embryogenesis. }

@@

%\begin{figure}[t]
%\includegraphics{fig1}
%\caption{Caricature illustrating the tasks of image registration and temporal ordering. {\xfigtextfontit (A)} Images of ``samples'', each in a different orientation and a different stage of development. {\xfigtextfontit (B)} Registered and ordered samples. For this caricature, the registration and ordering is straightforward because the data set is small, the landmarks are visually apparent, and the developmental changes are easy to recognize.}
%%\label{fig:fish}
%\customlabel{fig:fish}{1}
%\customlabel{subfig:fish_unordered}{\ref{fig:fish}{\it A}}
%\customlabel{subfig:fish_ordered}{\ref{fig:fish}{\it B}}
%\end{figure}

@@

%\begin{figure}[t]
%\includegraphics{fig2}
%\caption{Schematic illustrating angular synchronization and diffusion maps. {\xfigtextfontit (A)} Set of vectors, each in a different orientation. The pairwise alignment angles are indicated. {\xfigtextfontit (B)} The vectors from {\xfigtextfontit A}, each rotated about their midpoint so that the set is globally aligned. Note that the chosen rotation angles are consistent with the pairwise alignments in {\xfigtextfontit A}: the differences between a pair of angles in {\xfigtextfontit B} is the same as the pairwise angle in {\xfigtextfontit A}. {\xfigtextfontit (C)} Data points (in black) which lie on a one-dimensional nonlinear curve in two dimensions. Each pair of points is connected by an edge, and the edge weight is related to the Euclidean distance between the points through the diffusion kernel (see {\xfigtextfontit \SI}), so that close data points are connected by darker (``stronger'') edges. {\xfigtextfontit (D)} The data in {\xfigtextfontit C}, colored by the first (non-trivial) eigenvector from the diffusion map computational procedure. The color intensity is monotonic with the perceived curve arclength, thus parameterizing the curve.}
%%\label{fig:schematics}
%\customlabel{fig:schematics}{2}
%\customlabel{subfig:synch1}{\ref{fig:schematics}{\it A}}
%\customlabel{subfig:synch2}{\ref{fig:schematics}{\it B}}
%\customlabel{subfig:dmaps1}{\ref{fig:schematics}{\it C}}
%\customlabel{subfig:dmaps2}{\ref{fig:schematics}{\it D}}
%\end{figure}

%\begin{figure*}[t]
%\includegraphics{fig3}
%\caption{Method validation using live imaging of {\xfigtextfontit Drosophila} embryos.{\xfigtextfontit (A)} Selected images from a live imaging study of a {\xfigtextfontit Drosophilla} embryo during gastrulation. Each frame is in an arbitrary rotational orientation, and the order of the frames has been shuffled. {\xfigtextfontit (B)} Images from {\xfigtextfontit A} after registration and ordering using vector diffusion maps. The dorsal side of each embryo now appears at the top of each image, and the ventral side appears at the bottom. {\xfigtextfontit (C)} The correlation between the recovered rotation angle (using vector diffusion maps) and the true rotation angle {\xfigtextfontit (D)} The correlation between the recovered rank (using vector diffusion maps) and the true rank. {\xfigtextfontit (E)} The average error in the recovered angle and the rank correlation coefficient for $5$ different live imaging studies. {\xfigtextfontit (F)} The median rank correlation coefficient (over $200$ samples) as a function of the number of images in the data set. }
%\customlabel{fig:movie}{3}
%\customlabel{subfig:movie_unregistered_unordered}{\ref{fig:movie}{\it A}}
%\customlabel{subfig:movie_registered_ordered}{\ref{fig:movie}{\it B}}
%\customlabel{subfig:movie_angle_corr}{\ref{fig:movie}{\it C}}
%\customlabel{subfig:movie_rank_corr}{\ref{fig:movie}{\it D}}
%\customlabel{subfig:movie_error_table}{\ref{fig:movie}{\it E}}
%\customlabel{subfig:movie_bootstrap}{\ref{fig:movie}{\it F}}
%\end{figure*}

@@

%\begin{figure*}[t]
%\includegraphics{fig4}
%\caption{Analyzing images of fixed {\xfigtextfontit Drosophila} embryos. {\xfigtextfontit (A)} Images of {\xfigtextfontit Drosophila} embryos, stained for nuclei (gray), Twi (green), and dpERK (red). Each image is of a different embryo arrested at a different developmental time and in a different rotational orientation. {\xfigtextfontit (B)} Data from {\xfigtextfontit A}, registered and ordered using vector diffusion maps. The expert rank for each image is indicated. {\xfigtextfontit (C)} A representative ``developmental trajectory'' obtained from local averaging of the entire set of registered and ordered images (see \SI). {\xfigtextfontit (D)} Correlation between the image ranks calculated from the vector diffusion maps algorithm and the ranks obtained from ordering by an expert. The rank correlation coefficient is $0.97$. }
%\customlabel{fig:data2}{4}
%\customlabel{subfig:fixed_images_unregistered_unordered}{\ref{fig:data2}{\it A}}
%\customlabel{subfig:fixed_images_registered_ordered}{\ref{fig:data2}{\it B}}
%\customlabel{subfig:fixed_images_average_trajectory}{\ref{fig:data2}{\it C}}
%\customlabel{subfig:fixed_images_rank_corr}{\ref{fig:data2}{\it D}}
%\end{figure*}

@@

%\bibliographystyle{pnas2}
%\bibliography{../../references/references,../image_analysis_paper/background_reading/references}

\begin{thebibliography}{10}

\bibitem{jaeger2004dynamic}
Jaeger J, Surkova S, Blagov M, Janssens H, Kosman D, \textit{et~al.} (2004)
  Dynamic control of positional information in the early \textit{{D}rosophila}
  embryo. \textit{Nature} 430(6997):368--371.

\bibitem{peter2011gene}
Peter I~S, Davidson E~H (2011) A gene regulatory network controlling the
  embryonic specification of endoderm. \textit{Nature} 474(7353):635--639.

\bibitem{fowlkes2008quantitative}
Fowlkes C~C, Hendriks C~L~L, Ker{\"a}nen S~V, Weber G~H, R{\"u}bel O,
  \textit{et~al.} (2008) A quantitative spatiotemporal atlas of gene expression
  in the \textit{{D}rosophila} blastoderm. \textit{Cell} 133(2):364--374.

\bibitem{ng2012large}
Ng L~L, Sunkin S~M, Feng D, Lau C, Dang C, \textit{et~al.} (2012) Large-scale
  neuroinformatics for in situ hybridization data in the mouse brain.
  \textit{Int Rev Neurobiol} 104:159--182.

\bibitem{richardson2014emage}
Richardson L, Stevenson P, Venkataraman S, Yang Y, Burton N, \textit{et~al.}
  (2014) Emage: Electronic mouse atlas of gene expression, \textit{Mouse
  Molecular Embryology} (Springer), pp. 61--79.

\bibitem{castro2009automatic}
Castro C, Luengo-Oroz M, Desnoulez S, Duloquin L, Fernandez-de Manuel L,
  \textit{et~al.} (2009) An automatic quantification and registration strategy
  to create a gene expression atlas of zebrafish embryogenesis,
  \textit{Conf Proc IEEE Eng Med Biol Soc}, pp. 1469--1472.

\bibitem{zitova2003image}
Zitova B, Flusser J (2003) Image registration methods: a survey. \textit{Image
   Vis Comput} 21(11):977--1000.

\bibitem{rowley1998rotation}
Rowley H~A, Baluja S, Kanade T (1998) Rotation invariant neural network-based
  face detection, \textit{Conf Proc IEEE Comp Soc on Computer Vision and Pattern Recognition} (IEEE), pp. 38--44.

\bibitem{hajnal2010medical}
Hajnal J~V, Hill D~L (2010) \textit{Medical image registration} (CRC press).

\bibitem{greenspan1994rotation}
Greenspan H, Belongie S, Goodman R, Perona P (1994) Rotation invariant texture
  recognition using a steerable pyramid, \textit{Pattern Recognition, 1994.
  Vol. 2-Conference B: Computer Vision \& Image Processing., Proceedings of the
  12th IAPR International. Conference on} (IEEE), volume~2, pp. 162--167.

\bibitem{zhao2003face}
Zhao W, Chellappa R, Phillips P~J, Rosenfeld A (2003) Face recognition: A
  literature survey. \textit{ACM Computing Surveys (CSUR)} 35(4):399--458.

\bibitem{singer2012vector}
Singer A, Wu H~T (2012) Vector diffusion maps and the connection {L}aplacian.
  \textit{Commun Pure Appl Math} 65(8):1067--1144.

\bibitem{Belkin2003}
Belkin M, Niyogi P (2003) Laplacian eigenmaps for dimensionality reduction and
  data representation. \textit{Neural Comput} 15:1373--1396.

\bibitem{coifman2005geometric}
Coifman R~R, Lafon S, Lee A~B, Maggioni M, Nadler B, \textit{et~al.} (2005)
  Geometric diffusions as a tool for harmonic analysis and structure definition
  of data: Diffusion maps. \textit{Proc Natl Acad Sci USA} 102(21):7426--7431.

\bibitem{coifman2006geometric}
Coifman R~R, Lafon S (2006) Geometric harmonics: a novel tool for multiscale
  out-of-sample extension of empirical functions. \textit{Appl Comput Harmon Anal} 21(1):31--52.

\bibitem{tenenbaum2000global}
Tenenbaum J~B, De~Silva V, Langford J~C (2000) A global geometric framework for
  nonlinear dimensionality reduction. \textit{Science} 290(5500):2319--2323.

\bibitem{roweis2000nonlinear}
Roweis S~T, Saul L~K (2000) Nonlinear dimensionality reduction by locally
  linear embedding. \textit{Science} 290(5500):2323--2326.

\bibitem{zhao2014rotationally}
Zhao Z, Singer A (2014) Rotationally invariant image representation for viewing
  direction classification in cryo-em. \textit{J Struct Biol}
  186(1):153 -- 166.

\bibitem{singer2011viewing}
Singer A, Zhao Z, Shkolnisky Y, Hadani R (2011) Viewing angle classification of
  cryo-electron microscopy images using eigenvectors. \textit{SIAM J Imaging Sci} 4(2):723--759.

\bibitem{lafon2006data}
Lafon S, Keller Y, Coifman R~R (2006) Data fusion and multicue data matching by
  diffusion maps. \textit{IEEE Trans Pattern Anal Mach Intell} 28(11):1784--1797.

\bibitem{fernandez2014diffusion}
Fern{\'a}ndez {\'A}, Rabin N, Coifman R~R, Eckstein J (2014) Diffusion methods
  for aligning medical datasets: Location prediction in {CT} scan images.
  \textit{Med Image Anal} 18(2):425--432.

\bibitem{jaeger2012drosophila}
Jaeger J, Manu, Reinitz J (2012) \textit{{D}rosophila} blastoderm patterning.
  \textit{Curr Opin Genet Dev} 22(6):533 -- 541, genetics of system biology.

\bibitem{singer2011angular}
Singer A (2011) Angular synchronization by eigenvectors and semidefinite
  programming. \textit{Appl Comput Harmon Anal}
  30(1):20--36.

\bibitem{ahuja2007template}
Ahuja S, Kevrekidis I~G, Rowley C~W (2007) Template-based stabilization of
  relative equilibria in systems with continuous symmetry. \textit{Journal of
  Nonlinear Science} 17(2):109--143.

\bibitem{ian1998statistical}
Dryden I~L, Mardia K (1998) \textit{Statistical shape analysis}, Wiley series
  in probability and statistics: Probability and statistics (J. Wiley).

\bibitem{leptin2005gastrulation}
Leptin M (2005) Gastrulation movements: the logic and the nuts and bolts.
  \textit{Dev Cell} 8(3):305--320.

\bibitem{Lim2013kinetics}
Lim B, Samper N, Lu H, Rushlow C, Jim{\'e}nez G, \textit{et~al.} (2013)
  Kinetics of gene derepression by {ERK} signaling. \textit{Proc Natl Acad Sci USA} 110(25):10330--10335.

\bibitem{anavy2014blind}
Anavy L, Levin M, Khair S, Nakanishi N, Fernandez-Valverde S~L, \textit{et~al.}
  (2014) {BLIND} ordering of large-scale transcriptomic developmental
  timecourses. \textit{Development} 141(5):1161--1166.

\bibitem{trapnell2014dynamics}
Trapnell C, Cacchiarelli D, Grimsby J, Pokharel P, Li S, \textit{et~al.} (2014)
  The dynamics and regulators of cell fate decisions are revealed by
  pseudotemporal ordering of single cells. \textit{Nat Biotechnol}
  32(4):381--386.

\bibitem{gupta2008extracting}
Gupta A, Bar-Joseph Z (2008) Extracting dynamics from static cancer expression
  data. \textit{IEEE/ACM Trans Comput Biol Bioinform} 5(2):172--182.

\bibitem{yuan2014automated}
Yuan L, Pan C, Ji S, McCutchan M, Zhou Z~H, \textit{et~al.} (2014) Automated
  annotation of developmental stages of \textit{{D}rosophila} embryos in images
  containing spatial patterns of expression. \textit{Bioinformatics}
  30(2):266--273.

\bibitem{surkova2008characterization}
Surkova S, Kosman D, Kozlov K, Myasnikova E, Samsonova A~A, \textit{et~al.}
  (2008) Characterization of the \textit{{D}rosophila} segment determination
  morphome. \textit{Dev Biol} 313(2):844--862.

\bibitem{krzic2012multiview}
Krzic U, Gunther S, Saunders T~E, Streichan S~J, Hufnagel L (2012) Multiview
  light-sheet microscope for rapid in toto imaging. \textit{Nat Methods}
  9(7):730--733.

\bibitem{ichikawa2014live}
Ichikawa T, Nakazato K, Keller P~J, Kajiura-Kobayashi H, Stelzer E~H,
  \textit{et~al.} (2014) Live imaging and quantitative analysis of gastrulation
  in mouse embryos using light-sheet microscopy and 3d tracking tools.
  \textit{Nat Protoc} 9(3):575--585.

\bibitem{rubel2010coupling}
R{\"u}bel O, Ahern S, Bethel E, Biggin M~D, Childs H, \textit{et~al.} (2010)
  Coupling visualization and data analysis for knowledge discovery from
  multi-dimensional scientific data. \textit{Procedia Comput Sci}
  1(1):1757--1764.

\bibitem{chung2010microfluidic}
Chung K, Kim Y, Kanodia J~S, Gong E, Shvartsman S~Y, \textit{et~al.} (2011) A
  microfluidic array for large-scale ordering and orientation of embryos.
  \textit{Nat Methods} 8(2):171--176.

\end{thebibliography}
